---
title: "2D Moffat"
output:
  word_document: default
  html_document: default
date: "2024-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
library(plotly)
library(ggpointdensity)
library(viridis)
library(tidyr)
library(ggExtra)
library(ggpubr)
library(cowplot)
library(pracma)

## Color pallete
cbPalette <- c("#56B4E9", "#E69F00", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```
Importing the data and transforming for stages required at this analysis step.
```{r}
library(here)
i_am("2D Moffat.Rmd")
# laptop data from csv
Terzan5 <- readr::read_csv(here("Data", "Terzan 5 X-ray events.csv"), col_types = list(.default = readr::col_guess()), )
# head(Terzan5)

#fishbowl data from ODS
# library(readODS)
# Terzan5 <- read_ods(here("Raw Data.ods"), col_types = list(.default = readr::col_guess()), )
#removing extraneous columns
T5 <- data.frame(Terzan5$x, Terzan5$y)
colnames(T5) <- c('x','y')

#filtering into two Points of Interest
POI1 <- T5 %>%
  filter(x>4160 & x<4180 & y>4180 & y<4200)
POI2 <- T5 %>%
  filter(x>4105 & x< 4125 & y>4040 & y<4060)

```

The 2D Moffat Model has this form:

$$f(x,y) = A\left[1 + \frac{(x-x_{0})^2 + (y - y_{0})^2}{\alpha^2}\right]^{-\beta} $$
Where A = amplitude, $x0$ = x-position of the maximum, $y0$ = y-position of the maximum, $\alpha$ = core width, $\beta$ = power index.

Symmetric uncorr function - same influence

Similarly to the 1D Moffat, the amplitude must be greater than 0. Alpha can't be equal to zero and Beta must be greater than 0.

Visualising the 2D Moffat
```{r}
Moffat2D <- function(grid, parameters){
  x <- grid[[1]]
  y <- grid[[2]]
  
  amplitude <- parameters[[1]]
  x0 <- parameters[[2]]
  y0 <- parameters[[3]]
  alpha <- parameters[[4]]
  beta <- parameters[[5]]
  
  amplitude*(1+((x - x0)^2 + (y - y0)^2)/alpha^2)^(-beta)
}

#grid of variables
xRange <- seq(-5, 5, length = 100)
yRange <- seq(-5, 5, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

#Dummy parameters
dummyParams <- c(amplitude = 1, x0 = 0, y0 = 0, alpha = 1, beta = 1)

z <- Moffat2D(grid, dummyParams)
zMatrix <- matrix(z, nrow = length(xRange), ncol = length(yRange))

image(xRange, yRange, zMatrix, xlab = "x", ylab = "y", main = "2D Moffat")
persp(xRange, yRange, zMatrix, xlab = "x", ylab = "y", zlab = "f(x,y)", main = "2D Moffat")
```

Visualising the effects of changing parameters
```{r}
#baseline of 1,0,0,1,1
dummyParams <- c(amplitude = 1, x0 = 0, y0 = 0, alpha = 1, beta = 1)
dummyParams1 <- c(amplitude = 4, x0 = 0, y0 = 0, alpha = 1, beta = 1)
dummyParams2 <- c(amplitude = 1, x0 = 3, y0 = 0, alpha = 1, beta = 1)
dummyParams3 <- c(amplitude = 1, x0 = 0, y0 = -2, alpha = 1, beta = 1)
dummyParams4 <- c(amplitude = 1, x0 = 0, y0 = 0, alpha = 3, beta = 1)
dummyParams5 <- c(amplitude = 1, x0 = 0, y0 = 0, alpha = 1, beta = 2.6)

#baseline plot
z <- Moffat2D(grid, dummyParams)
zMatrix <- matrix(z, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, zMatrix, xlab = "x", ylab = "y", main = "2D Moffat")

#increasing amplitude
z1 <- Moffat2D(grid, dummyParams1)
z1Matrix <- matrix(z, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, z1Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Increasing A")

#moving x0
z2 <- Moffat2D(grid, dummyParams2)
z2Matrix <- matrix(z2, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, z2Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Moving x0")

#moving y0
z3 <- Moffat2D(grid, dummyParams3)
z3Matrix <- matrix(z3, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, z3Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Moving y0")

#Increasing alpha
z4 <- Moffat2D(grid, dummyParams4)
z4Matrix <- matrix(z4, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, z4Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Increasing alpha")

#Increasing beta
z5<- Moffat2D(grid, dummyParams5)
z5Matrix <- matrix(z5, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, z5Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Increasing beta")

persp(xRange, yRange, zMatrix, xlab = "x", ylab = "y", main = "2D Moffat")
persp(xRange, yRange, z1Matrix, xlab = "x", ylab = "y", main = "2D Moffat - Increasing A")
```

The Negative Log Likelihood of the 2D Moffat Model:
$$-\sum(\ln\left[A\right]-\beta\ln\left[1+\frac{(x-x_0)^2 + (y-y_0)^2}{\alpha^2}\right]) $$
```{r}
NLL <- function(grid, parameters){
  predictedDens <- Moffat2D(grid, parameters)
  return(-sum(log(predictedDens)))
}
```


Investigating how the parameters effect the negative log likelihood function
```{r}
#changing the value of the amplitude
amplitude <- seq(0.25, 5, by = 0.25)

i <- 1
changingA <- vector("numeric", length = 20)

while(i<=20){
  parameters <- c(amplitude[[i]], 1, 1, 1, 1)
  grid <- c(0,0)
  changingA[i] <- (NLL(grid, parameters))
  i <- i+1
}

#changing the value of x0
x0 <- seq(-5, 5, by = 0.5)

i <- 1
changingx0 <- vector("numeric", length = 21)

while(i<=21){
  parameters <- c(1, x0[[i]], 1, 1, 1)
  grid <- c(0,0)
  changingx0[i] <- (NLL(grid, parameters))
  i <- i+1
}

#changing the value of y0
y0 <- seq(-5, 5, by = 0.5)

i <- 1
changingy0 <- vector("numeric", length = 21)

while(i<=21){
  parameters <- c(1, 1, y0[[i]], 1, 1)
  grid <- c(0,0)
  changingy0[i] <- (NLL(grid, parameters))
  i <- i+1
}

#changing the value of alpha
alpha <- seq(0.25, 10, by = 0.25)

i <- 1
changingAlpha <- vector("numeric", length = 40)

while(i<=40){
  parameters <- c(1, 1, 1, alpha[[i]], 1)
  grid <- c(0,0)
  changingAlpha[i] <- (NLL(grid, parameters))
  i <- i+1
}

#changing the value of beta
beta <- seq(0.25, 10, by = 0.25)

i <- 1
changingBeta <- vector("numeric", length = 40)

while(i<=40){
  parameters <- c(1, 1, 1, 1, beta[[i]])
  grid <- c(0,0)
  changingBeta[i] <- (NLL(grid, parameters))
  i <- i+1
}

par(mfrow = c(2,3))
plot(amplitude, changingA, type = "l", xlab = "Amplitude", ylab = "Negative Log Likelihood")
plot(x0, changingx0, type = "l", xlab = "x0", ylab = "Negative Log Likelihood")
plot(y0, changingy0, type = "l", xlab = "x0", ylab = "Negative Log Likelihood")
plot(alpha, changingAlpha, type = "l", xlab = "Alpha", ylab = "Negative Log Likelihood")
plot(beta, changingBeta, type = "l", xlab = "Beta", ylab = "Negative Log Likelihood")
```
This is similar to the plots of the effects the parameters of the 1D Moffat Function have on the Negative Log Likelihood.

Amplitude needs to be expressed in terms of alpha and beta.

$$f(x,y) = \frac{\beta-1}{\pi\alpha^2}\left[1 + \frac{(x-x_{0})^2 + (y - y_{0})^2}{\alpha^2}\right]^{-\beta} $$
Validating the 2D Moffat PDF
```{r}
Moffat2DPDF <- function(parameters, x, y){
  x0 <- parameters[[1]]
  y0 <- parameters[[2]]
  alpha <- parameters[[3]]
  beta <- parameters[[4]]
  
  alpha <- pmax(alpha, 1e-6)
  beta <- pmax(beta, 1 + 1e-6)
  
  return(((beta-1)/(pi*alpha^2))*(1+((x - x0)^2 + (y - y0)^2)/alpha^2)^(-beta))
}

library(pracma)
#Test 1
T1Params <- c(x0 = 1, y0 = 4, alpha = 0.4, beta = 2)
MoffatIntegral2D1 <- integral2(Moffat2DPDF, xmin = -100, xmax = 100, ymin = -100, ymax = 100, parameters = T1Params)[[1]]
MoffatIntegral2D1

#Test 2
T2Params <- c(x0 = -3, y0 = 7, alpha = 1.2, beta = 1.3)
MoffatIntegral2D2 <- integral2(Moffat2DPDF, xmin = -100, xmax = 100, ymin = -100, ymax = 100, parameters = T2Params)[[1]]
MoffatIntegral2D2

#Test 3
T3Params <- c(x0 = 4, y0 = 0, alpha = 3, beta = 2.4)
MoffatIntegral2D3 <- integral2(Moffat2DPDF, xmin = -100, xmax = 100, ymin = -100, ymax = 100, parameters = T3Params)[[1]]
MoffatIntegral2D3
```
```{r}
xRange <- seq(-20, 20, length = 200)
yRange <- seq(-20, 20, length = 200)
grid <- expand.grid(xRange, yRange)

A1 <- (T1Params[[4]]-1)/(pi * T1Params[[3]])
zT1 <- Moffat2D(grid, parameters = c(A1, T1Params))
zT1Matrix <- matrix(zT1, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, zT1Matrix)

A2 <- (T2Params[[4]]-1)/(pi * T2Params[[3]])
zT2 <- Moffat2D(grid, parameters = c(A2, T2Params))
zT2Matrix <- matrix(zT2, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, zT2Matrix)

A3 <- (T3Params[[4]]-1)/(pi * T3Params[[3]])
zT3 <- Moffat2D(grid, parameters = c(A3, T3Params))
zT3Matrix <- matrix(zT3, nrow = length(xRange), ncol = length(yRange))
image(xRange, yRange, zT3Matrix)

```

Gradient of the negative log likelihood with respect to each parameter
$$\frac{\delta}{\delta x_0} = -\sum\frac{2\beta(x-x_0)}{\alpha^2+(x-x_0)^2+(y-y_0)^2}$$
$$\frac{\delta}{\delta y_0} = -\sum\frac{2\beta(y-y_0)}{\alpha^2+(x-x_0)^2+(y-y_0)^2}$$
$$ \frac{\delta}{\delta\alpha} = 2\sum(\frac{1}{\alpha} + \frac{\beta((x-x_0)^2 + (y-y_0)^2)}{\alpha^3 + \alpha((x-x_0)^2+(y-y_0)^2)})$$
$$ \frac{\delta}{\delta\beta} = -\sum(\frac{1}{\beta} - \ln\left[1+\frac{(x-x_0)^2 + (y-y_0)^2}{\alpha^2}\right])$$
```{r}
NLL2D <- function(parameters, x, y){
  predictedDens <- Moffat2DPDF(parameters, x, y)
  return(-sum(log(predictedDens)))
}
```

```{r}
deriv2DNLL <- function(parameters, x, y){
  x0 <- parameters[[1]]
  y0 <- parameters[[2]]
  alpha <- parameters[[3]]
  beta <- parameters[[4]]
  
 dx0 <- -sum((2*beta*(x-x0))/(alpha^2 + (x-x0)^2 + (y-y0)^2))
 dy0 <- -sum((2*beta*(y-y0))/(alpha^2 + (x-x0)^2 + (y-y0)^2))
 dAlpha <- 2*sum(1/alpha + (beta*((x-x0)^2 + (y-y0)^2))/(alpha^3 + alpha*((x-x0)^2 + (y-y0)^2)))
 dBeta <- -sum(1/beta - log(1+((x-x0)^2 + (y-y0)^2)/alpha^2))
  
  return(c(dx0, dy0, dAlpha, dBeta))
}
```

```{r}
mean(POI1$x)
mean(POI1$y)
```
```{r}
initGuess <- c(x0 = mean(POI1$x), y0 = mean(POI1$y), alpha = 1, beta = 2)
lowerBounds <- c(0, 0, 1e-6, 1+1e-6)

resultPOI1 <- optim(par = initGuess, fn = NLL2D, gr = deriv2DNLL, x = POI1$x, y = POI1$y, method = "L-BFGS-B", lower = lowerBounds)
resultPOI1
optimPOI1Par <- resultPOI1$par
optimPOI1Par
```
Viewing the optimised function and viewing the data.
```{r}
xRange <- seq(4160, 4180, length = 100)
yRange <- seq(4180, 4200, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

amplitude <- (optimPOI1Par[[4]]-1)/(pi*optimPOI1Par[[3]]^2)
parameters <- c(amplitude, optimPOI1Par)

zPOI1<- Moffat2D(grid, parameters)
zPOI1Matrix <- matrix(zPOI1, nrow = length(xRange), ncol = length(yRange))
par(pty = "s")
image(xRange, yRange, zPOI1Matrix, xlab = "x", ylab = "y", main = "2D Moffat")

model <- ggplot(POI1, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000)) + 
  coord_fixed()

contours <- ggplot(grid, aes(x = x, y = y)) + 
  stat_contour(aes(z = zPOI1), bins = 10) +
  coord_fixed()

ggMarginal(model, type = "histogram")
contours
```
```{r}
contourData <- ggplot_build(contours)$data[[1]] #extracting the contours (from the optimised function)
combined <- model + geom_path(data = contourData, aes(x = x, y = y, group = group), color = "red", alpha = 0.5)
combined
```
Zooming in on this point of interest to try to get a clearer picture.
```{r}
library(ggExtra)
#refiltering the POI1
POI1ZOOM <- POI1 %>%
  filter(x>4167 & x<4177 & y>4183 & y<4193)

#creating a point density plot
modelZoomedInPOI1 <- ggplot(POI1ZOOM, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000)) + 
  coord_fixed()

contourData <- ggplot_build(contours)$data[[1]]

#combining the contour plot and the point density plot
combined <- modelZoomedInPOI1 + geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[5], alpha = 1, linewidth= 0.3)

#viewing
ggMarginal(combined, type = "histogram")

#plotting contours on the raw data
ggplot(POI1ZOOM, aes(x,y)) +
  geom_point(size = 0.2) + 
  geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[2], alpha = 1) +
  coord_fixed()
```
Add the marginal PDF of the model on top of the histogram

```{r}
mean(POI2$x)
mean(POI2$y)
```


```{r}
initGuess <- c(x0 = 4115.227, y0 = 4052.706, alpha = 1, beta = 2)

result2D <- optim(par = initGuess, fn = NLL2D, gr = deriv2DNLL, x = POI2$x, y = POI2$y, method = "L-BFGS-B", lower = lowerBounds)
result2D
optimPOI2Par <- result2D$par
optimPOI2Par

```
```{r}
#check, the area under the curve should be equal to 1.
integral2(Moffat2DPDF, xmin = 3000, xmax = 5000, ymin = 3000, ymax = 5000, parameters = optimPOI2Par)[[1]]
#maximum amplitude
Moffat2DPDF(optimPOI2Par, 4115.26, 4051.90)
```

```{r}
xRange <- seq(4110, 4120, length = 100)
yRange <- seq(4045, 4055, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

amplitude <- (optimPOI2Par[[4]]-1)/(pi*optimPOI2Par[[3]]^2)
#amplitude
parameters <- c(amplitude, optimPOI2Par)

zPOI2<- Moffat2D(grid, parameters)
zPOI2Matrix <- matrix(zPOI2, nrow = length(xRange), ncol = length(yRange))
par(pty = "s")
image(xRange, yRange, zPOI2Matrix, xlab = "x", ylab = "y", main = "2D Moffat")
```

```{r}
xRange <- seq(4110, 4120, length = 100)
yRange <- seq(4045, 4055, length = 100)
data <- expand.grid(x = xRange, y = yRange)

model <- ggplot(POI2, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000)) +
  theme(legend.position = "none")

contours <- ggplot(data, aes(x = x, y = y)) + 
  stat_contour(aes(z = zPOI2), bins = 10) 

ggMarginal(model, type = "histogram")
contours
```
```{r}
contourData <- ggplot_build(contours)$data[[1]]

combined <- model + geom_path(data = contourData, aes(x = x, y = y, group = group), color = "red", alpha = 0.5)

combined
```
Zooming in on this point of interest to try to get a clearer picture.
```{r}
#refiltering the POI2
POI2ZOOM <- POI2 %>%
  filter(x>4110 & x<4120 & y>4048 & y<4058)

#creating a point density plot
modelZoomedInPOI2 <- ggplot(POI2ZOOM, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000)) + 
  coord_fixed()

contourData <- ggplot_build(contours)$data[[1]]

#combining the contour plot and the point density plot
combined <- modelZoomedInPOI2 + geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[5], alpha = 1, linewidth= 0.3)

#viewing
ggMarginal(combined, type = "histogram")

#plotting contours on the raw data
ggplot(POI2ZOOM, aes(x,y)) +
  geom_point(size = 0.2) + 
  geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[2], alpha = 1) +
  coord_fixed()
```
Alternate zoom method
```{r}
combined + coord_cartesian(xlim = c(4110, 4120), ylim = c(4048, 4058))
```

I would like to display the Marginal PDF's of the model.
$$f_X(x) = \int_{-\infty}^{\infty} f_{XY}(x,y) dy$$
$$f_Y(y) = \int_{-\infty}^{\infty} f_{XY}(x,y) dx$$

Obtaining Marginal PDF's - Technique 1
```{r}
xRange <- seq(4105, 4125, length = 200)
yRange <- seq(4040, 4060, length = 200)
data <- expand.grid(x = xRange, y = yRange)

#for all values within the above grid, apply the 2D Moffat PDF function with the optimised parameters. This will generate a grid of values. Where X is rows and Y are columns
jointPDFValues <- outer(X = xRange, Y = yRange, Moffat2DPDF, parameters = optimPOI2Par)

#check
tot <- sum(jointPDFValues)

#To obtain the marginal PDFS, I need to integrate the joint function with respect to one of the variables. 
#Instead I sum up all the values in each row. This has the effect of integrating with respect to y. As I take the value from each position over the range. 
marginalX <- apply(jointPDFValues, 1, sum)
marginalY <- apply(jointPDFValues, 2, sum)

data.frame(xRange, marginalX)
#checks - the sum of each marginal distribution should equal the sum of the whole area.
sum(marginalX)
sum(marginalY)

plot(xRange, marginalX, type = "l", xlab = "x", ylab = "Density", main = "Marginal PDF of x")
plot(yRange, marginalY, type = "l", xlab = "y", ylab = "Density", main = "Marginal PDF of y")
```
Obtaining Marginal PDFs - Technique 2
```{r}
xRange <- seq(4105, 4125, length = 200)
yRange <- seq(4040, 4060, length = 200)

Rx <- length(xRange) - 1
Ry <- length(yRange) - 1

marginalX <- numeric(length = Rx)
for(i in 1:Rx){
  marginalX[i] <- integral2(Moffat2DPDF, xmin = xRange[i], xmax = xRange[i+1], ymin = 4040, ymax = 4060, parameters = optimPOI2Par)[[1]]
}

sum(marginalX) #checking it sums to 1 - expect it to be a little less as it isn't summed over infinite (x,y) plain

marginalY <- numeric(length = Ry)
for(i in 1:Ry){
  marginalY[i] <- integral2(Moffat2DPDF, xmin = 4105, xmax = 4125, ymin = yRange[i], ymax = yRange[i+1], parameters = optimPOI2Par)[[1]]
}

sum(marginalY) #checking it sums to 1

xMids <- numeric(length = Rx)
for(i in 1:Rx){
  xMids[i] <- (xRange[i]+xRange[i+1])/2
}
yMids <- numeric(length = Ry)
for(i in 1:Ry){
  yMids[i] <- (yRange[i]+yRange[i+1])/2
}

dataframeMarginalX <- data.frame(x = xMids, y = marginalX)
dataframeMarginalY <- data.frame(x = yMids, y = marginalY)

plot(dataframeMarginalX, type = "l")
plot(dataframeMarginalY, type = "l")

```

Plotting Marginal PDFs and histograms from the model
```{r}
margX <- ggplot(dataframeMarginalX) +
  geom_line(aes(x = xMids, y = marginalX), col = cbPalette[2]) +
  labs(x = "x", y = "Density", main = "Marginal PDF of x") 

margY <- ggplot(dataframeMarginalY) +
  geom_line(aes(x = yMids, y = marginalY), col = cbPalette[2]) +
  labs(x = "y", y = "Density", main = "Marginal PDF of y") +
  rotate()

histX <- ggplot(POI2, aes(x = x)) +
  geom_histogram(aes(y = after_stat(density)), bins = 100, col = cbPalette[3], fill = cbPalette[5])

histY <- ggplot(POI2, aes(x = y)) +
  geom_histogram(aes(y = after_stat(density)), bins = 100, col = cbPalette[3], fill = cbPalette[5]) + 
  rotate()

margX
margY
histX
histY
```
```{r}
xRange <- seq(4105, 4125, length = 200)
yRange <- seq(4040, 4060, length = 200)
xMarg <- numeric(length = 200)
for(i in 1:200){
  xMarg[i] <- Moffat2DPDF(optimPOI2Par, xRange[i], optimPOI2Par[2])
}

xMargdf <- data.frame(xRange, xMarg)

histogramWithMarginalX <- ggplot(POI2, aes(x = x)) +
  geom_histogram(aes(y = after_stat(density)), bins = 250, col = cbPalette[3], fill = cbPalette[5]) +
  geom_line(data = xMargdf, aes(x = xRange, y = xMarg), col = cbPalette[2]) +
  labs(x = NULL, y = "Density") + 
  theme(axis.text.x = element_blank(), axis.title.x = element_blank())

histogramWithMarginalX

yMarg <- numeric(length = 200)
for(i in 1:200){
  yMarg[i] <- Moffat2DPDF(optimPOI2Par, optimPOI2Par[1], yRange[i])
}

yMargdf <- data.frame(yRange, yMarg)

histogramWithMarginalY <- ggplot(POI2, aes(x = y)) +
  geom_histogram(aes(y = after_stat(density)), bins = 250, col = cbPalette[3], fill = cbPalette[5]) +
  geom_line(data = yMargdf, aes(x = yRange, y = yMarg), col = cbPalette[2]) +
  labs(x = NULL , y = "Density") + 
  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
  rotate()
        
histogramWithMarginalY

plot_grid(histogramWithMarginalX, NULL, combined, histogramWithMarginalY, ncol = 2, align = "hv", rel_widths = c(2, 1), rel_heights = c(1, 2), axis = c("l", "b"))
```

```{r}
celestialSource1 <- T5 %>%
  filter(x>4030 & x<4050 & y>4060 & y<4080)
celestialSource2 <- T5 %>%
  filter(x>4070 & x<4080 & y>4100 & y<4110)
celestialSource3 <- T5 %>%
  filter(x>4070 & x<4080 & y>4150 & y<4160)
```

```{r}
ggplot(celestialSource1, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(5, 25, 125)) + 
  coord_fixed()
ggplot(celestialSource2, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(5, 25, 125)) + 
  coord_fixed()
ggplot(celestialSource3, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(5, 25, 125)) + 
  coord_fixed()
```


Looking at other sources
Celestial Source 1
```{r}
mean(celestialSource1$x)
mean(celestialSource1$y)
```
```{r}
initGuess <- c(x0 = 4039.865, y0 = 4071.066, alpha = 1, beta = 2)
lowerBounds <- c(0, 0, 1e-6, 1+1e-6)

result2D <- optim(par = initGuess, fn = NLL2D, gr = deriv2DNLL, x = celestialSource1$x, y = celestialSource1$y, method = "L-BFGS-B", lower = lowerBounds)
optimCS1Par <- result2D$par
optimCS1Par
```
```{r}
xRange <- seq(4030, 4050, length = 100)
yRange <- seq(4060, 4080, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

amplitude <- (optimCS1Par[[4]]-1)/(pi*optimCS1Par[[3]]^2)
parameters <- c(amplitude, optimCS1Par)

zCS1<- Moffat2D(grid, parameters)
zCS1Matrix <- matrix(zCS1, nrow = length(xRange), ncol = length(yRange))
par(pty = "s")
image(xRange, yRange, zCS1Matrix, xlab = "x", ylab = "y", main = "2D Moffat")
```
```{r}
xRange <- seq(4030, 4050, length = 100)
yRange <- seq(4060, 4080, length = 100)
datac1 <- expand.grid(x = xRange, y = yRange)

model <- ggplot(celestialSource1, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(4, 16, 64)) + 
  coord_fixed()

contours <- ggplot(datac1, aes(x = x, y = y)) + 
  stat_contour(aes(z = zCS1), bins = 10) +
  coord_fixed()

ggMarginal(model, type = "histogram")
contours
```
```{r}
contourData <- ggplot_build(contours)$data[[1]]
combined <- model + geom_path(data = contourData, aes(x = x, y = y, group = group), color = "red", alpha = 0.5)
combined
```
```{r}
#refiltering Celestial Source 1
CS1ZOOM <- celestialSource1 %>%
  filter(x>4035 & x<4045 & y>4070 & y<4080)

#creating a point density plot
modelZoomedInCS1 <- ggplot(CS1ZOOM, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000)) + 
  coord_fixed()

contourData <- ggplot_build(contours)$data[[1]]

#combining the contour plot and the point density plot
combined <- modelZoomedInCS1 + geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[5], alpha = 1, linewidth= 0.3)

#viewing
ggMarginal(combined, type = "histogram")

#plotting contours on the raw data
ggplot(CS1ZOOM, aes(x,y)) +
  geom_point(size = 0.2) + 
  geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[2], alpha = 1) +
  coord_fixed()
```

Celestial Source 2
```{r}
mean(celestialSource2$x)
mean(celestialSource2$y)
```

```{r}
initGuess <- c(x0 = 4073.925, y0 = 4105.563, alpha = 1, beta = 2)
lowerBounds <- c(0, 0, 1e-6, 1+1e-6)

result2D <- optim(par = initGuess, fn = NLL2D, gr = deriv2DNLL, x = celestialSource2$x, y = celestialSource2$y, method = "L-BFGS-B", lower = lowerBounds)
optimCS2Par <- result2D$par
optimCS2Par
```
```{r}
xRange <- seq(4070, 4080, length = 100)
yRange <- seq(4100, 4110, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

amplitude <- (optimCS2Par[[4]]-1)/(pi*optimCS1Par[[3]]^2)
parameters <- c(amplitude, optimCS2Par)

zCS2<- Moffat2D(grid, parameters)
zCS2Matrix <- matrix(zCS2, nrow = length(xRange), ncol = length(yRange))
par(pty = "s")
image(xRange, yRange, zCS2Matrix, xlab = "x", ylab = "y", main = "2D Moffat")
```
```{r}
xRange <- seq(4070, 4080, length = 100)
yRange <- seq(4100, 4110, length = 100)
datacs2 <- expand.grid(x = xRange, y = yRange)

model <- ggplot(celestialSource2, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(5, 25, 125)) + 
  coord_fixed()

contours <- ggplot(datacs2, aes(x = x, y = y)) + 
  stat_contour(aes(z = zCS2), bins = 10) +
  coord_fixed()

ggMarginal(model, type = "histogram")
contours
```
```{r}
contourData <- ggplot_build(contours)$data[[1]]
combined <- model + geom_path(data = contourData, aes(x = x, y = y, group = group), color = "red", alpha = 0.5)
combined
```
```{r}
#plotting contours on the raw data
ggplot(celestialSource2, aes(x,y)) +
  geom_point(size = 0.2) + 
  geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[2], alpha = 1) +
  coord_fixed()
```


Celestial Source 3
```{r}
mean(celestialSource3$x)
mean(celestialSource3$y)
```
```{r}
initGuess <- c(x0 = 4074.631, y0 = 4156.233, alpha = 1, beta = 2)
lowerBounds <- c(0, 0, 1e-6, 1+1e-6)

result2D <- optim(par = initGuess, fn = NLL2D, gr = deriv2DNLL, x = celestialSource3$x, y = celestialSource3$y, method = "L-BFGS-B", lower = lowerBounds)
optimCS3Par <- result2D$par
optimCS3Par
```
```{r}
xRange <- seq(4070, 4080, length = 100)
yRange <- seq(4150, 4160, length = 100)
grid <- expand.grid(x = xRange, y = yRange)

amplitude <- (optimCS3Par[[4]]-1)/(pi*optimCS1Par[[3]]^2)
parameters <- c(amplitude, optimCS3Par)

zCS3<- Moffat2D(grid, parameters)
zCS3Matrix <- matrix(zCS3, nrow = length(xRange), ncol = length(yRange))
par(pty = "s")
image(xRange, yRange, zCS3Matrix, xlab = "x", ylab = "y", main = "2D Moffat")
```
```{r}
xRange <- seq(4070, 4080, length = 100)
yRange <- seq(4150, 4160, length = 100)
datac1 <- expand.grid(x = xRange, y = yRange)

model <- ggplot(celestialSource3, aes(x,y)) +
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(5, 25, 125)) + 
  coord_fixed()

contours <- ggplot(datac1, aes(x = x, y = y)) + 
  stat_contour(aes(z = zCS3), bins = 10) +
  coord_fixed()

ggMarginal(model, type = "histogram")
contours
```
```{r}
contourData <- ggplot_build(contours)$data[[1]]
combined <- model + geom_path(data = contourData, aes(x = x, y = y, group = group), color = "red", alpha = 0.5)
combined
```
```{r}
#viewing
ggMarginal(combined, type = "histogram")

#plotting contours on the raw data
ggplot(celestialSource3, aes(x,y)) +
  geom_point(size = 0.2) + 
  geom_path(data = contourData, aes(x = x, y = y, group = group), color = cbPalette[2], alpha = 1) +
  coord_fixed()
```

Modelled Areas
```{r}
ggplot(T5,aes(x,y)) + 
  geom_pointdensity(adjust = 0.05, size = 0.1, shape = "1") + 
  scale_color_viridis(direction = -1, option = "B", trans = "log", breaks = c(10, 100, 1000, 10000)) + 
  coord_fixed() + 
  #POI1
  geom_segment(aes(x = 4160, y = 4180, xend = 4180, yend = 4180)) +
  geom_segment(aes(x = 4160, y = 4200, xend = 4180, yend = 4200)) +
  geom_segment(aes(x = 4160, y = 4180, xend = 4160, yend = 4200)) +
  geom_segment(aes(x = 4180, y = 4180, xend = 4180, yend = 4200)) + 
  #POI2
  geom_segment(aes(x = 4105, y = 4040, xend = 4125, yend = 4040)) +
  geom_segment(aes(x = 4105, y = 4060, xend = 4125, yend = 4060)) +
  geom_segment(aes(x = 4105, y = 4040, xend = 4105, yend = 4060)) +
  geom_segment(aes(x = 4125, y = 4040, xend = 4125, yend = 4060)) + 
  #CS1
  geom_segment(aes(x = 4030, y = 4060, xend = 4050, yend = 4060)) +
  geom_segment(aes(x = 4030, y = 4080, xend = 4050, yend = 4080)) +
  geom_segment(aes(x = 4030, y = 4060, xend = 4030, yend = 4080)) +
  geom_segment(aes(x = 4050, y = 4060, xend = 4050, yend = 4080)) + 
  #CS2
  geom_segment(aes(x = 4070, y = 4100, xend = 4080, yend = 4100)) +
  geom_segment(aes(x = 4070, y = 4110, xend = 4080, yend = 4110)) +
  geom_segment(aes(x = 4070, y = 4100, xend = 4070, yend = 4110)) +
  geom_segment(aes(x = 4080, y = 4100, xend = 4080, yend = 4110)) + 
  #CS3
  geom_segment(aes(x = 4070, y = 4150, xend = 4080, yend = 4150)) +
  geom_segment(aes(x = 4070, y = 4160, xend = 4080, yend = 4160)) +
  geom_segment(aes(x = 4070, y = 4150, xend = 4070, yend = 4160)) +
  geom_segment(aes(x = 4080, y = 4150, xend = 4080, yend = 4160))
```

