---
title: "Moffat1D"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2024-03-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(plotly)
library(readr)
library(ggpointdensity)
library(viridis)
library(tidyr)
library(fitdistrplus)
```
Importing the data and narrowing it down to a "1D" slice of a Celestial Source
Viewing some of those slices
```{r}
library(here)
i_am("MoffatPDF.Rmd")
Tz5 <- readr::read_csv(here("Data", "Terzan 5 X-ray events.csv"), col_types = list(.default = readr::col_guess()), )
head(Tz5)
T5 <- data.frame(Tz5$x, Tz5$y)
colnames(T5) <- c('x','y')
POI1 <- T5 %>%
  filter(x>4160 & x<4180 & y>4180 & y<4200)
POI2 <- T5 %>%
  filter(x>4105 & x< 4125 & y>4040 & y<4060)
#data to fit to
line2 <- POI2 %>%
  filter(y>4050.5 & y< 4051.5)
plotdist(line2$x)
plot(line2)

line2.5 <- line2 %>%
  filter(y>4050.5 & y< 4051)
plot(line2.5)

xl2 <- line2$x
xl2.5 <- line2.5$x
xVall2 <- seq(min(xl2), max(xl2), length = 963)
xVall2.5 <- seq(min(xl2.5), max(xl2.5), length = 291)
```

This Rmarkdown document is to create a function of the 1D Moffat distribution function where the amplitude is given in respect to the parameters that describe the core width and power index.

The equation is as follows:

$$f(x) = \frac{\Gamma(\beta)}{\alpha\sqrt\pi~\Gamma(\beta-\frac{1}{2})} \left[1+\frac{(x-\mu)^2}{\alpha^2}\right]^{-\beta}$$
In order for the amplitude to remain positive, alpha must be greater than 0 and beta must be greater than or equal to 1.5.

```{r}
MOFFATPDF <- function(parameters, x){
  mu <- parameters[[1]]
  alpha <- parameters[[2]]
  beta <- parameters[[3]]
  
  alpha <- pmax(alpha, 1e-6)
  beta <- pmax(beta, 0.5+1e-6)
  
  predictedDensity <- (gamma(beta)/(alpha*sqrt(pi)*gamma(beta-0.5)))*(1+((x-mu)/alpha)^2)^-beta
  
}
```

To use the optim function likes to minimize functions.
For that I need a Negative Log Likelihood

```{r}
NLL <- function(parameters, x){
  predictedDensity <- MOFFATPDF(parameters, x)
  return(-sum(log(predictedDensity)))
}
```

Now I need the partial derivatives of the negative log likelihood with respect to each parameter

$$-\sum\ln\left[\Gamma(\beta)\right] - \ln\left[\alpha\sqrt\pi~\Gamma(\beta-\frac{1}{2})\right] - \beta\ln\left[1+\frac{(x-\mu)^2}{\alpha^2}\right]$$
$$\frac{\delta}{\delta\mu} = -\sum\frac{2\beta(x-\mu)}{\alpha^2 + (x-\mu)^2}$$
$$\frac{\delta}{\delta\alpha} = -\sum-\frac{1}{\alpha} + \frac{2\beta(x-\mu)^2}{\alpha^3 + \alpha(x-\mu)^2}$$
To obtain the partial derivative with respect to beta I will be using the Rcode of digamma
$$\frac{\delta}{\delta\beta} = -\sum digamma(\beta) - digamma(\beta - \frac{1}{2}) - \ln\left[1+\frac{(x-mu)^2}{alpha^2}\right]$$

```{r}
derivMOFFATPDS <- function(parameters, x){
  mu <- parameters[[1]]
  alpha <- parameters[[2]]
  beta <- parameters[[3]]
  
  
  dMu <- -sum((2*beta*(x-mu))/(alpha^2 + (x-mu)^2))
  dAlpha <- -sum(-1/alpha + (2*beta*(x-mu)^2)/(alpha^3+alpha*(x-mu)^2))
  dBeta <- -sum(digamma(beta)-digamma(beta-0.5)-log(1+(x-mu)^2/alpha^2))
  
  return(c(dMu, dAlpha, dBeta))
}
```

Trying with optim
```{r}
initialGUESS <- c(mu = 4115, alpha = 1, beta = 4)
lowerBOUNDS <- c(0, 1e-6, 1.5)

RESULT <- optim(par = initialGUESS, fn = NLL, gr = derivMOFFATPDS, x = xl2.5, method = "L-BFGS-B", lower = lowerBOUNDS)
optimPAR2.5 <- RESULT$par
optimPAR2.5

optimDENS2.5 <- MOFFATPDF(optimPAR2.5, xVall2.5)
plot(xVall2.5, optimDENS2.5, type = "l", main = "Optimised Density Function", xlab = "x", ylab = "Density")
```
```{r}
RESULT <- optim(par = initialGUESS, fn = NLL, gr = derivMOFFATPDS, x = xl2, method = "L-BFGS-B", lower = lowerBOUNDS)
optimPAR2 <- RESULT$par
optimPAR2

optimDENS2 <- MOFFATPDF(optimPAR2, xVall2)
plot(xVall2, optimDENS2, type = "l", main = "Optimised Density Function", xlab = "x", ylab = "Density")
```

