---
title: "1D Moffat"
output: html_document
date: "2024-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown is focused on fitting the 1D Moffat Model to the slices of individual celestial sources.

```{r}
library(ggplot2)
library(plotly)
library(readr)
library(ggpointdensity)
library(viridis)
library(tidyr)
library(fitdistrplus)
```

```{r}
library(here)
here()
Tz5 <- readr::read_csv(here("Data", "Terzan 5 X-ray events.csv"), col_types = list(.default = readr::col_guess()), )
head(Tz5)
```


Downloading the data and sorting it into two celestial sources
```{r}
T5 <- data.frame(Tz5$x, Tz5$y)
colnames(T5) <- c('x','y')
POI1 <- T5 %>%
  filter(x>4160 & x<4180 & y>4180 & y<4200)
POI2 <- T5 %>%
  filter(x>4105 & x< 4125 & y>4040 & y<4060)
```


Functions required for the optim function
```{r}
Moffat <- function(parameters, x){

  amplitude <- parameters[[1]]
  mu <- parameters[[2]]
  gamma <- parameters[[3]]
  alpha <- parameters[[4]]

  amplitude <- pmax(amplitude, 1e-6)
  gamma <- pmax(gamma, 1e-6)
  alpha <- pmax(alpha, 1e-6)

  predictedDensity <- amplitude*(1+((x-mu)^2/gamma^2))^(-alpha)
}

negLogLike <- function(parameters, x){
  predictedDens <- Moffat(parameters, x)
  return(-sum(log(predictedDens)))
}

MoffatDeriv <- function(parameters, x){
  amplitude <- parameters[[1]]
  mu <- parameters[[2]]
  gamma <- parameters[[3]]
  alpha <- parameters[[4]]
  
  amplitude <- pmax(amplitude, 1e-6)
  gamma <- pmax(gamma, 1e-6)
  alpha <- pmax(alpha, 1e-6)

  d_A = -sum(1/amplitude)
  d_mu = sum((2*alpha*(x - mu))/(gamma^2 + (x - mu)^2))
  d_gamma = sum((2*alpha*(x - mu)^2)/(gamma^3 + gamma*(x - mu)^2))
  d_alpha = sum(log(1+((x-mu)^2)/gamma^2))
  
  return(c(d_A, d_mu, d_gamma, d_alpha))
}

```


Slicing the individual celestial sources to create '1D' lines

```{r}
#data to fit to
line2 <- POI2 %>%
  filter(y>4050.5 & y< 4051.5)
plotdist(line2$x)
plot(line2)

xl2 <- line2$x
plot(xl2)

line2.5 <- line2 %>%
  filter(y>4050.5 & y< 4051)
plot(line2.5)
xl2.5 <- line2.5$x
plot(xl2.5)
```
Using the Optim Function
```{r}
#set initial guesses
initialGuess <- c(amplitude = 0.9, mu = 2, gamma = 1, alpha = 0.5)
#set bounds
lowerBounds <- c(1e-6, -Inf, 1e-6, 1e-6)

#sampledata
sampleData <- rnorm(1000, mean = 2, sd = 1)

result <- optim(par = initialGuess, fn = negLogLike, gr = MoffatDeriv, x = sampleData, method = "L-BFGS-B", lower = lowerBounds)
optimPar <- result$par
print(optimPar)
```

Fitting to the data
```{r}
xVall2 <- seq(min(xl2), max(xl2), length = 963)

#set initial guesses
initialGuess <- c(amplitude = 1, mu = 4115, gamma = 1, alpha = 1)

#Method 2
#l2m2 = line 2 (slice of celestial source), method 2 (mle)
resultl2m2 <- optim(par = initialGuess, fn = negLogLike, gr = MoffatDeriv, x = xl2, lower = lowerBounds, method = "L-BFGS-B")
optimParl2m2 <- resultl2m2$par
print(optimParl2m2)
optDenl2m2<- Moffat(optimParl2m2, xVall2)
```
```{r}
plot(xVall2, optDenl2m2, type = "l", main = "Optimised Density Function", xlab = "x", ylab = "Density")

par(mfrow = c(1,2))
plot(xl2)
plot(Moffat(optimParl2m2, xl2))
```

Halving the previous slice
```{r}
#Fitting to data
xVall2.5 <- seq(min(xl2.5), max(xl2.5), length = 963)

#set initial guesses
initialGuess <- c(amplitude = 1, mu = 4115, gamma = 1, alpha = 1)

#l2.5m2 = line 2.5 (slice of celestial source), method 2 (mle)
resultl2.5m2 <- optim(par = initialGuess, fn = negLogLike, gr = MoffatDeriv, x = xl2.5, lower = lowerBounds, method = "L-BFGS-B")
optimParl2.5m2 <- resultl2.5m2$par
print(optimParl2.5m2)
optDenl2.5m2<- Moffat(optimParl2.5m2, xVall2.5)
```

Plotting the function with the optimised parameters
```{r}
plot(xVall2.5, optDenl2.5m2, type = "l", main = "Optimised Density Function", xlab = "x", ylab = "Density")

par(mfrow = c(1,2))
plot(xl2.5)
plot(Moffat(optimParl2.5m2, xl2.5))
```

